---
title: "Visualization_1"
output: github_document
---

```{r}
library(tidyverse)
library(ggridges)
library(p8105.datasets)

data("weather_df")
weather_df
```

## Basic Scatterplot

Only create blank canvas:
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax))
```
Add `geom_point`
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```
Easier just pipe in:
```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```
Save the plot by assigning it to an object:
Call it again to put it from "Value" to "Data" at "Environment
```{r}
ggp_weather = 
  weather_df |>
  ggplot(aes(x = tmin, y = tmax)) 

ggp_weather + geom_point()

```
Check that some rows are missing:
```{r}
weather_df |>
  filter(is.na(tmax))
```


## Advanced Scatterplot

`geom_point`adding color point based on variable group
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name))
```

Define aesthetics can matter:
`alpha` to show different density 
`geom_smooth` to add smooth curve fits the scatters, but not very trustworthy
`geom_smooth(se = FALSE)`to remove susceptible error ban 
`color = name` give different color based on variable `name`
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  geom_smooth(se = FALSE)
```

Use faceting real quick:
`facet_grid`to split into multiple plots by subgroups:
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```
`facet_grid(name ~ .)` to show each subgroup in rows
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(name ~ .)
```
`x = date` change x variable to shows how y changes in the time of year
`size = prcp` shows information about percipertation
```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name, shape = name)) + 
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

Learning Assignment Code:
Write a code chain that starts with weather_df; focuses only on Central Park, converts temperatures to Fahrenheit, makes a scatterplot of min vs. max temperature, and overlays a linear regression line (using options in geom_smooth()).

`geom_smooth(method()` Smoothing method (function) to use, accepts either NULL or a character vector, e.g. "lm", "glm", "gam", "loess" or a function
`geom_smooth(se = )` Display confidence interval around smooth? (TRUE by default, see `level` to control.)
```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  mutate(
    tmax_fahr = tmax * (9 / 5) + 32,
    tmin_fahr = tmin * (9 / 5) + 32) |> 
  ggplot(aes(x = tmin_fahr, y = tmax_fahr)) +
  geom_point(alpha = .5) + 
  geom_smooth(method = "lm", se = FALSE)
```

## Small things

you can also remove the points, or change the order of`point` and `smooth` depend on the order you put
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name, shape = name)) +
 # geom_point(alpha = .5) + 
  geom_smooth(se = FALSE)
```

Show density of how data clusters:
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name, shape = name)) +
  geom_hex()
```
Color the points:
`color` worked for both `geom_point()` and `geom_smooth()`, but `shape` only applies to `geom_point`
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name, shape = name)) +
  geom_point(color = "purple")
```

## Odds and ends

No `geom_point`, instead using `geom_smooth` directly
```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_smooth(se = FALSE) 
```

`geom_hex()`, `geom_bin2d()`, or `geom_density2d()` all can use to avoid overplotting
```{r}
ggplot(weather_df, aes(x = tmax, y = tmin)) + 
  geom_hex()
```

Learning Assignment Code:
```{r}
ggplot(weather_df) + geom_point(aes(x = tmax, y = tmin), color = "blue")
ggplot(weather_df) + geom_point(aes(x = tmax, y = tmin, color = "blue"))
```


## Univariate Plots

Making histogram, no need to mention y-axis, it auto counts.
```{r}
ggplot(weather_df, aes(x = tmax)) + 
  geom_histogram()
```
`color` the bound, and `fill` color the area inside the bound:
```{r}
ggplot(weather_df, aes(x = tmax)) + 
  geom_histogram(color = "purple", fill = "red")
```

Stack and differnetiate subgroups by color
```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_histogram(position = "dodge", binwidth = 2)
```

Density plot:
```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue")
```

Box plot: `x = name` separate by subgroup variable - name
```{r}
ggplot(weather_df, aes(x = name, y = tmax)) + 
  geom_boxplot(aes(fill = name))
```

Violin Plots:(vertical version of density plot)
```{r}
ggplot(weather_df, aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), alpha = .5) + 
  stat_summary(fun = "median", color = "blue")
```
Ridge Plot: nice if you have lots of categories in which the shape of the distribution matters
```{r}
ggplot(weather_df, aes(x = tmax, y = name)) + 
  geom_density_ridges(scale = .85)
```


## Show a ZOOM-IN verison of plot

Filter out data in the end (long tails):
```{r}
weather_df |>
  filter(prcp > 5, prcp < 1000) |>
  ggplot (aes(x = prcp, fill = name)) +
  geom_density(alpha = 0.2)
```


## Saving and embedding plots

```{r}
ggp_weather = 
  ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) 

ggsave("ggp_weather.pdf", ggp_weather, width = 8, height = 5)
```

Embedding plots in an R Markdown document:
`fig.width`, `fig.height`, and `fig.asp` control the size of the figure created by R
`out.width` or `out.heigh` control size of the figure inserted into your document
```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

Font and point size in the figures generated by R are constant, they don’t scale with the overall size of the figure.
*Text in a figure with width 12 will look smaller than text in a figure with width 6 after both have been embedded in a document
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name))
```



